\section{Bug hunting}

\subsection{Compiling for \szadb{}}
Lets have a closer look at the C program below.  Its stated purpose is to
replace, if possible, runs of white space with tabs.  The width of a tab
is fixed and equal to a constant \name{TABSTOP}.  If a text line is 
getting too long then substitutions are abandoned and remaining
characters are copied without modifications.
\bigskip
\listing{\exmpskip}{unexpand.c}
\bigskip
\noindent
This program has actually two bugs.
See if you can find them just examinig the source.

Here is how \szadb\ can help. Compile source as follows
(these commands are for Sozobon~C)
\begin{exmpl}
	{\tt cc -t -O -o unexpand.tos unexpand.c}
\end{exmpl}
Flag \name{-O} is not necessary but with this particular compiler
you will probably find a disassembled code easier to follow.
You may test the compiled executable on its own source.
\begin{exmpl}
	{\tt unexpand.tos <unexpand.c >output }
\end{exmpl}
In the first moment the program appears to work, but a closer examination
of the \name{output} reveals that the indentation is not exactly right.
Moreover, some lines start with a blank, followed by a tab, which is
not really what was intended.
There are also other problems.  Check yourself.
The likely suspect will be an array \name{tabs} of tabstops filled by
a function \name{settab()}.

\subsection{The first bug}
Start the program under \szadb\ control
\begin{exmpl}
	{\tt adb.ttp unexpand.tos}
\end{exmpl}
and set a breakpoint at \verb|settab+4|, just after {\tt link} instruction.
Run the program with \verb|:c ; $C|. This will produce the following
display
\readexample{bcktrace.exm}
In the absence of better information all arguments 
shown in the stack backtrace are assumed to be two bytes wide.  
We know from the source that \name{settab()} actualy
expects one pointer.
Confirm that it got a right one by putting it together from two halfs
and using the request \verb|68bb4=p| to print it as the symbol.
You should see \verb|_tabs| in response.
To continue execution of the current function 
till it returns to its caller, use {\tt :f} which stands for {\tt :finish}.

It is clear from lines 30 and 35 that the first tabstop in
\verb|_tabs| is expected to be on a position \hbox{{\tt TABSTOP - 1}}.
Dumping some initial fragment of the just initialized array 
we see the following:
\readexample{tabs.exm}
This is clearly wrong and one bug becomes obvious.
To repair it line 59 should be changed to
\begin{exmpl}
	\verb|for (i = 1; i <= MAXLIN; i++) {...}|
\end{exmpl}

\subsection{\ldots{}and the other one}

The second bug is harder to spot, since for most of test inputs our
program will work correctly.
This is a typical example of a program broken for boundary conditions.
To make it easier to track the problem set
\name{MAXLIN} to some small integer (around 10 should be good),
recompile the program, restart the debugging session and set a breakpoint
at \verb|main+28|, just after a character was read.
Set this breakpoint with a count and a request 
to show a received character with
\begin{exmpl}
	{\tt main+28,5:b <d0=cx}
\end{exmpl}
With carefuly chosen input this will show where you are in the
program and will skip unnecessary stops in the same time.
You have to provide an input by typing it yourself.
Remember that to repeat the last command  {\tt :c} it is enough to hit
\key{Return}.
For execution defaults \szadb\ will use the most recently typed command
even if some other requests were executed by breakpoints.

In order to see how the received character is processed you can single step
with \name{:s}, which will follow program execution.
The request \name{:n} will single step like \name{:s} but will
execute function calls at full speed and so have the
effect of stepping over them. Breakpoints set on skipped
levels still will be obeyed.

Tracing some tests inputs with the value of the variable \name{col} around
\name{MAXLIN} should reveal the second error soon enough.  If you
want to try it yourself, do not read further.

Looking at line 17 we find a sharp less-than inequality in the test. It should
be replaced by a less-than-equal-to. Otherwise, when the value of col
equals MAXLIN, and your input is "right", the program is trying to read, in
lines 28 and 31, from the location \verb|&tabs[col]|,
which is one past the end of the array.
The remaining analysis of this bug is left as an exercise to the reader.


\subsection{Breaking out}

Another stepping command is the \name{:j} jump request.  Its purpose
it to short-circuit loops.  It behaves exactly like \name{:n} in that
it skips-over function calls but unlike the \name{:n} request which
follows branch instructions, the \name{:j} will place a temporary
breakpoint at the instruction immediately following the current
instruction in memory.  Think of \name{:n} as step-next-logical
instruction and \name{:j} as step-next-physical instruction.  The idea
is that you only use \name{:j} to step-out of loops where you are
sitting on the loop-back branch.
\readexample{jumpone.exm}
In the example above, if you are sitting on the \name{bnz} instruction and you
wish to execute the remainder of the loop at full speed then you use the
\name{:j}, which places a temporary breakpoint at the location \name{.end}.
However care
must be taken when using this command because it is possible that the 
flow-of-control never reaches the temporary breakpoint.
\readexample{jumptwo.exm}
This is a case where NOT to use the \name{:j} command. 
Basically to use the \name{:j}
request you should know how your C compiler sets up its loops or where your
assembler code is meant to go. It is sometime a good idea to put a safety
breakpoint somewhere you know you will end up.

Because of the unusual nature of the \name{:j} request,
it will not autorepeat if the next command is just a \key{return} key. 
Instead \name{:n} is performed. Also it is
also considered a variant of \name{:n} when attaching execution requests to
stepping commands (see further down).

It should be also mentioned that all stepping commands have upper case
counterparts, which are noisier. When used they return the a full register
display after the step (\name{:S} is like doing \verb|:s;$r|).
