%%%%%%%%%%%%%%%%%%%%%%%
%   szadb primer      %
%   version 1.3       %
%   August, 1990      %
%%%%%%%%%%%%%%%%%%%%%%%
\documentstyle[twoside,11pt,verblist]{article}

\title{\protect{\LARGE{\tt szadb} {\bf primer}}}
\author{Micha{\l} Jaegermann \\[.6in]
	edited by\\
	Anthony Howe}
\date{13 Sept 1990}

%%
\newcommand{\szadb}{{\tt szadb}}
\newcommand{\adb}{{\tt adb.ttp}}
\newcommand{\name}[1]{{\tt #1}}
\newcommand{\key}[1]{{\sf $\langle$#1$\rangle$}}
\newcommand{\carret}{\char94}
\newlength{\exmpskip}
\setlength{\exmpskip}{2\parindent}
\newenvironment{exmpl}
  {\begin{quote}\setlength{\leftskip}{\parindent}}{\end{quote}}
\newcommand{\readexample}[1]
   {\medskip\inputverbatim[\exmpskip]{#1}\medskip\noindent}
%%
%\date{June 4, 1990}
\begin{document}
\vfill
\maketitle
\thispagestyle{empty}
%%
\titlepage\null\vfill{\sf 
\begin{quote}
	There is NO WARRANTY with respect to this publication, or the
	program it describes, and disclaim any implied or explicit
	suggestions of usefulness for any particular purpose.  Use
	this program only if you are willing to assume all risks, and
	damages, if any, arising as a result, even if caused by
	negligence or other fault.

\bigskip
	Sozobon C compiler and \szadb\ debugger are both copyright
	\copyright 1989 by Sozobon Ltd.  Both can be freely copied
	and distributed provided all copyright notices will remain
	intact and all modified versions will be clearly marked as such.

\medskip
	Atari, ST, TOS and GEMDOS are trademarks of Atari Co.\\
	GEM is a trademark of Digital Research Co.\\
	Mark Williams C is a trademark or Mark Williams Co.\\
	{\sc Unix} is a trademark of AT\&T.
\end{quote}
}\par\null\endtitlepage
\sloppy
%{\tt  % turn this on if you want to make life easy for dvi2tty
%\raggedright
\section{Introduction}

\szadb\ is an assembly level debugger for programs written for the TOS 
operating system on the Atari ST computer.
It is primarily meant to be a companion
to the freely distributed Sozobon C compiler, and can be copied and passed
around under the same conditions.

One of design goals of the original \szadb\ project was to provide 
a debugger closely resembling {\tt adb},
which is available on most {\sc Unix} systems. There are some dissimilarities
which follow mostly from differences in target machines
and operating systems. Still users with an {\tt adb} experience should feel
right at home.

The originally released version 1.0 was written by Johann Ruegg and Don Dugger.
This document describes version 1.2 prepared by Anthony Howe and
Micha{\l} Jaegermann.

This version provides multi-command input lines, which can be attached to
breakpoints and stepping commands. It gives an opportunity to record a
debugging sesion in a file, to define function keys, and other assorted
niceties. It accepts the symbol table formats of both Sozobon C (Alcyon C style)
and Mark Williams C.

If your compiler's symbol table format is not supported, or there are changes
or additions you wish to add, the source is provided free with the debugger.

\input{firsteps}
\input{running}
\input{debug}
\input{topics}
%}   % turn this on if you want to make life easy for dvi2tty
\end{document}
